---
title: "Experiment_with_data"
author: "Saurabh Gupta"
date: "June 12, 2017"
output: pdf_document
---

## Problem statement

In this problem, we have to classify the "Income.Group" based on the predictors. We use rpart classifiers to classify the dataset.

### Load the library

```{r}
suppressMessages(library(caret))
suppressMessages(library(doParallel))
suppressMessages(library(Hmisc))
library(data.table)

```

### Data loading and exploration

We load the data from given urls. Our goal is to predict the Income.Group for testData and we prepare our model on Vehicle dataset.

```{r}
cl <- makeCluster(detectCores())
registerDoParallel(cl)

#fileUrlTrain <- "https://datahack-prod.s3.ap-south-1.amazonaws.com/workshop_train_file/train_gbW7HTd.csv"
#download.file(fileUrlTrain,destfile = "./fileTrain.csv")
Vehicle <- read.csv("fileTrain.csv",header=T,na.strings = "")

#fileUrlTest <- "https://datahack-prod.s3.ap-south-1.amazonaws.com/workshop_test_file/test_2AFBew7.csv"
#download.file(fileUrlTest,destfile = "./fileTest.csv")
testData <- read.csv("fileTest.csv",header = T,na.strings = "")

```

### Data Cleaning

We convert the Income.Group value ( if it is "<=50K" it gives X0 otherwise it gives X1)

```{r}
names(Vehicle) <- gsub("[.]","_",names(Vehicle))
names(testData) <- gsub("[.]","_",names(testData))
Vehicle$Income_Group <- factor(ifelse(Vehicle$Income_Group=="<=50K",0,1))
Vehicle$Income_Group <- make.names(Vehicle$Income_Group)
Vehicle$Income_Group <- factor(Vehicle$Income_Group)
```

### Data Visualization

Classes in the response variable is unbalanced. Hence sampling is need to get the  good accuracy.

```{r}
str(Vehicle)
table(Vehicle$Income_Group)  
prop.table(table(Vehicle$Income_Group))
```

### Missing value detection and treatment

We replaced missing value to the value which occurs frequently in that predictor

```{r}
colSums(is.na(Vehicle))
Vehicle$Workclass <- impute(Vehicle$Workclass,mode)
Vehicle$Occupation <- impute(Vehicle$Occupation,mode)
Vehicle$Native_Country <- impute(Vehicle$Native_Country,mode)
```


### Create train and test data
```{r}
index <- createDataPartition(Vehicle$Income_Group,p=0.7,list=FALSE)
training <- Vehicle[index,]
testing <- Vehicle[-index,]
```


### Model Building

```{r}
ctrl <- trainControl(method="repeatedcv",
                     repeats = 10,
                     classProbs = T,
                     sampling = "up",
                     allowParallel = TRUE,
                     summaryFunction = twoClassSummary
)

rpart_mod <- train(Income_Group ~ .,
                   data=training,
                   method="rpart",
                   trControl=ctrl,
                   tuneLength=30,
                   metric="ROC",
                   na.action = na.omit
)


```


### Prediction for training data

```{r}
predicted <- predict(rpart_mod,testing)
caret::confusionMatrix(predicted,testing$Income_Group)
```


### ROC 

```{r}
suppressMessages(library(ROSE))
roc.curve(testing$Income_Group,predicted)
```

### Prediction for testData

```{r}
# predicted <- ifelse(predicted == "X0","<=50K",">50K")
# df <- data.frame(ID = testData$ID,Income.Group=predicted)
# write.csv(df,file = "final_solutions.csv")
```



